Follow-up Samples

### 2017-04-10 submission/manuscript.Rmd
Editing with Pat

---
title: ''
csl: microbiome.csl
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  word_document: default
geometry: margin=1.0in
bibliography: reference.bib
---

@@

$2$ Department of Internal Medicine, University of Michigan, Ann Arbor, MI

$3$ Department of Family Medicine and Community Medicine, Penn State Hershey Medical Center, Hershey, PA

\end{center}

@@

## Abstract

**Background:** Colorectal cancer (CRC) is a worldwide health problem and research suggests a correlation between the fecal bacterial microbiome and CRC.  Despite this, very little is known about what happens to the microbiome after treatment for an adenoma or carcinoma.  This study tested the hypothesis that treatment for adenoma or carcinoma results in changes towards a normal bacterial community.  Specifically, we tried to identify components within the community that were different before and after treatment of adenoma, advanced adenoma, and carcinoma.

**Results:** There was a larger change in the bacterial community in response to treatment for carcinoma versus adenoma (P-value `r ifelse(change_summary["thetayc", "Pvalue_adnVcrc"] < 0.05, invisible("<"), invisible(">"))` 0.05) but not carcinoma versus advanced adenoma (P-value `r ifelse(change_summary["thetayc", "Pvalue_srnVcrc"] < 0.05, invisible("<"), invisible(">"))` 0.05).  There was a trend for increasingly less community similarity, between samples pre- and post-treatment from adenoma to advanced adenoma to carcinoma.  Despite this, no difference was found in the relative abundance of any specific OTU before and after treatment for adenoma, advanced adenoma, or carcinoma groups (P-value `r ifelse(min(all_OTU_pair_wilcox$BH_corrected, na.rm = TRUE) < 0.05, invisible("<"), invisible(">"))` 0.05).  Using Random Forest models to assess whether changes in post-treatment samples were towards a normal community, only those with carcinoma had a significant decrease in positive probability (P-value `r ifelse(all_probs_paired_wilcox["carcinoma", "BH_correction"] < 0.05, invisible("<"), invisible(">"))` 0.05); providing further evidence that treatment has the greatest effect in those with carcinoma. The adenoma model used a total of `r length(rownames(adn_imp_vars))` OTUs, the SRN model used a total of `r length(rownames(srn_imp_vars))` OTUs, and the carcinoma model used a total of `r length(rownames(crc_imp_vars))` OTUs.  A total of `r length(common_vars_summary$otu)` OTUs were common to all three models with many classifying to commensal bacteria (e.g. *Lachnospiraceae*, *Bacteroides*, *Anaerostipes*, *Blautia*, and *Dorea*).  Both chemotherapy and radiation did not provide any additional changes to the bacterial community in those treated for carcinoma (P-value `r ifelse(min(chemo_rad_summary$bh, na.rm = TRUE) < 0.05, invisible("<"), invisible(">"))` 0.05).

**Conclusions:** Our data partially supports the hypothesis that the microbiome changes after treatment towards a normal community.  Individuals with carcinoma had more drastic differences to the overall community than those with adenoma.  Common OTUs to all models were overwhelmingly from commensal bacteria, suggesting that these bacteria may be important in initial polyp formation, development of advanced adenoma, and transition to carcinoma.

@@

bacterial microbiome; colorectal cancer; polyps; FIT; post-surgery; risk factors

@@

Colorectal cancer (CRC) is currently the third most common cause of cancer deaths [@jemal_cancer_2010; @haggar_colorectal_2009].  The rate of disease mortality has seen a significant decrease, thanks mainly to improvements in screening [@jemal_cancer_2010].  However, despite this improvement there are still approximately 50,000 deaths from the disease per year [@haggar_colorectal_2009].  Some of these deaths are due to disease recurrence, with current estimates that 20-30% of those undergoing resection having a CRC recurrence [@hellinger_reoperation_2006].  This is important since approximately 30-50% of those who have a recurrence will die of CRC [@ryuk_predictive_2014].  To reduce the burden on the healthcare system, finding ways to better stratify those at highest risk of recurrence is of great importance.

Recent studies in humans and mouse models have shown that altered membership and structure of the gut microbiome  correlate with CRC pathogenesis [@zackular_manipulation_2016; @arthur_microbial_2014; @zackular_gut_2013].  Further, bacterial communities on the mucosa are altered between normal and tumor tissue [@dejea_microbiota_2014].    Collectively, these studies provide a tantalizing link between our gut bacteria and CRC and suggest the possibility of using biomarkers to diagnose disease.  Indeed, models created with microbiome data and clinical tests, such as Fecal Immunoglobulin Test (FIT), result in good predictions of CRC [@baxter_microbiota-based_2016; @zeller_potential_2014].  While these studies show how changed microbial communities or invasion by inflammatory species can impact CRC progression  [@flynn_metabolic_2016], they provide very little information as to whether these communities change and rebound towards normal after successful treatment of adenoma, advanced adenoma, or carcinoma.

@@

Using pre- (initial) and post- (follow up) treatment samples we tested the hypothesis that treatment causes detectable changes towards a normal microbiome in those with adenoma, advanced adenoma, and carincoma.  First, we assessed differences between samples pre- and post-treatment in adenoma, advanced adenoma, or carcinoma using alpha or beta diversity metrics.  Second, we explored whether models built to classify adenoma, advanced adenoma, or carcinoma versus normal were able to identify specific community members that differed between pre- and post-treatment samples.  We also used these models to assess whether changes in the community were toward a more normal microbiome.  Finally, we assessed both whether surgery for adenomas and advanced adenomas provided larger community changes and whether chemotherapy or radiation provided additional changes to the microbiome over surgical resection. This study aims to provide evidence as to whether the altered CRC microbiome persists or shifts back towards normal after such interventions.

@@

***The Bacterial Community:*** 
Within our `r length(rownames(good_metaf))`-person cohort we tested whether the microbiome in patients with adenoma (n = `r change_summary["thetayc", "adn_n"]`), advanced adenoma (n = `r change_summary["thetayc", "srn_n"]`), or carcinoma (n = `r change_summary["thetayc", "crc_n"]`) had any broad differences between pre- and post-treatment samples.  We found that carcinoma patients had a more dissimilar bacterial community between these samples than those with adenoma (P-value `r ifelse(change_summary["thetayc", "Pvalue_adnVcrc"] < 0.001, invisible("<"), invisible(">"))` 0.001) [Figure 1A].  Although no significant differences were observed between advanced adenoma and carcinoma there was an increase in the dissimilarity of pre- and post-treatment samples from adenoma (`r paste(format(change_summary["thetayc", "adn_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "adn_SD"], digits = 2))` (mean `r paste(label = "\u00B1")` SD)) to advanced adenoma (`r paste(format(change_summary["thetayc", "srn_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "srn_SD"], digits = 2))`) to carcinoma (`r paste(format(change_summary["thetayc", "crc_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "crc_SD"], digits = 2))`) [Figure 1A].  The bacterial community structure pre- post-treatment was visualized using NMDS for adenoma [Figure 1B] (PERMANOVA `r ifelse(beta_summary["Adenoma", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05), advanced adenoma [Figure 1C] (PERMANOVA `r ifelse(beta_summary["SRN", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05), and carcinoma [Figure 1D] (PERMANOVA `r ifelse(beta_summary["Carcinoma", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05).  Interestingly, when pre- and post-treatment samples were compared, regardless of whether they were adenoma or carcinoma, there was no significant overall difference in beta diversity (PERMANOVA `r ifelse(beta_summary["All", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05).  There was no difference between pre- and post-treatment samples when investigating alpha diversity metrics for adenoma, advanced adenoma, or carcinoma [Table S1].  Additionally, there was also no difference in the relative abundance of any specific OTU between these samples for adenoma, advanced adenoma, or carcinoma only [Figure S1].

@@

***Common OTUs:***
Next, to identify which OTUs were important at all stages of disease, we identified common predictive OTUs within the adenoma, advanced adenoma, and carcinoma models.  When we compared the three different models with each other there were a total of `r length(rownames(common_vars_summary))` common OTUs.  Some of the most common taxonomic identifications belonged to *Bacteroides*, *Blautia*, *Anaerostipes*, *Lachnospiraceae*, and *Dorea*.  The vast majority of the OTUs that were common between these models had classifications to bacteria typically thought of as commensal [Table S2]. The rank of these OTUs, based on median mean decrease in accuracy (MDA), varied higly depending on whether they were in the adenoma, advanced adenoma, or carcinoma model [Figure 3].  As an example, *Bacteroides* (OTU85) was ranked `r filter(crc_varsMDA, otu == "Otu000085") %>% select(rank)`^th^ in the carcinoma model but in both the adenoma and advanced adenoma model ranked `r filter(adn_varsMDA, otu == "Otu000085") %>% select(rank)`^th^ and `r filter(srn_varsMDA, otu == "Otu000085") %>% select(rank)`^st^ respectively [Figure 3].

@@

Many of the common OTUs that we identified taxonomically classified to potential butyrate producers (e.g. *Clostridiales*, *Roseburia*,  and *Anaerostipes*) [Table S2].  Other OTUs classified to bacteria that are inhibited by polyphenols (e.g. *Bacteroides*). Both butyrate and polyphenols are thought to be protective against cancer, in part by reducing inflammation [@louis_gut_2014].  These protective compounds are derived from the breakdown of fiber, fruits, and vegetables by resident gut microbes. One example of this potential diet-microbiome-inflammation-polyp axis is that *Bacteroides*, which was highly prevalent in our models, are known to be increased in those with high non-meat based protein consumption [@zhu_intake_2016].  High protein consumption in general has been linked with an increased CRC risk [@mu_colonic_2016].  Conversely, *Bacteroides* are inhibited by polyphenols which are derived from fruits and vegetables [@ozdal_reciprocal_2016].  Our data fits with the hypothesis that the microbial metabolites from breakdown products within our own diet could not only help to shape the existing community but also have an effect on CRC risk and disease progression.  Within this context the commensal community may be an important modifiable risk factor for monitoring and preventing CRC recurrence after treatment.

@@

Building off of our results, an area for future research, is that within our study we do not know whether individuals who were still classified as positive by the carcinoma model eventually had a subsequent CRC recurrence. This information would help to strengthen the case for this model keeping numerous individuals above the cutoff threshold even though at follow up they were diagnosed as no longer having carcinoma.  It would also provide additional evidence that the microbiome could be used as a risk stratification tool in monitoring recurrence risk and whether different interventions could potentially change this community and lower the probability of future reccurence.

@@

**Figure 1: General Differences between Adenoma, Advanced Adenoma, and Carcinoma Groups After Treatment.**  A) A significant difference was found between the adenoma and carcinoma group for thetayc (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`).  B) NMDS of the pre- and post-treatment samples for the adenoma group.  C) NMDS of the pre- and post-treatment samples for the advanced adenoma group. D) NMDS of the pre- and post-treatment samples for the carcinoma group.

@@

**Figure S2: ROC Curves of the Adenoma, Advanced Adenoma, and Carcinoma Models.**  A) Adenoma ROC curve: The light green shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark green line represents the model using 100% of the data set and what was used for subsequent classification.  B) Advanced Adenoma ROC curve: The light yellow shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark yellow line represents the model using 100% of the data set and what was used for subsequent classification.  C) Carcinoma ROC curve: The light red shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark red line represents the model using 100% of the data set and what was used for subsequent classification.

**Figure S3: Summary of Important Variables for the Adenoma, Advanced Adenoma, and Carcinoma Models.**  A) MDA of the most important variables in the adenoma model.  The dark green point represents the mean and the lighter green points are the value of each of the 100 different runs.  B) Summary of Important Variables in the advanced adenoma model.  MDA of the most important variables in the SRN model.  The dark yellow point represents the mean and the lighter yellow points are the value of each of the 100 different runs.  C) MDA of the most important variables in the carcinoma model.  The dark red point represents the mean and the lighter red points are the value of each of the 100 different runs.

>>>>

---
title: ''
csl: microbiome.csl
fontsize: 12pt
output:
  pdf_document:
. includes:
. . in_header: header.tex
. keep_tex: yes
  word_document: default
geometry: margin=1.0in
bibliography: reference.bib
---

@@

$2$ Department of Internal Medicine, University of Michigan, Ann Arbor, M.

$3$ Department of Family Medicine and Community Medicine, Penn State Hershey Medical Center, Hershey, P.

\end{center.

@@

## Abstract [ edit at end ]

**Background.** Colorectal cancer (CRC) is a worldwide health problem. Despite growing evidence that members of the gut microbiota can drive tumorigenesis, little is known about what happens to the microbiome after treatment for an adenoma or carcinoma. This study tested the hypothesis that treatment for adenoma or carcinoma alters the abundance of bacterial populations associated with with disease to those associated with a normal colon. We tested this hypothesis by sequencing the 16S rRNA genes in the feces of `r length(rownames(good_metaf))` individuals before and after treatment for adenoma (N=`r change_summary["thetayc", "adn_n"]`), advanced adenoma (N=`r change_summary["thetayc", "srn_n"]`), and carcinoma (N=`r change_summary["thetayc", "crc_n"]`).

**Results.** There were large changes the the bacterial communities associated with treatment across the three groups. The communities from patients with carcinomas changed significantly more than those with adenoma following treatment (P-value=`r change_summary["thetayc", "Pvalue_adnVcrc"]`); however, there was not a significant difference between those with advanced adenoma and those with adenoma or carcinoma (P-value>0.05). Although treatment brought about large intrapersonal changes, the change in the abundance of individual OTUs to treatment was not consistent within diagnosis groups (P-value=`r ifelse(min(all_OTU_pair_wilcox$BH_corrected, na.rm = TRUE) < 0.05, invisible("<"), invisible(">"))` 0.05). Because the distribution of OTUs across patients and diagnosis groups was patchy, we used the Random Forest machine learning algorithm to identify groups of OTUs that allowed us to successfully distinguish between pre- and post-treatment samples for each of the diagnosis groups. The OTUs that were shared across the models for the three diagnosis groups were dominated by lineages that are commonly considered to be commensals (e.g. *Lachnospiraceae*, *Bacteroides*, *Anaerostipes*, *Blautia*, and *Dorea*).

4. Models suggest that the positive predictive value from diagnostic models decrease following treatment of carcinomas

Finally, we used a larger cohort that contained individuals with

Using diagnostic Random Forest models that were trained to differentiate between individuals with normal colons or those with adenomas, advanced adenomas, or carcinomas we measured the change in the positive probability of having one of the three diagnoses.

Using Random Forest models to assess whether changes in post-treatment samples were towards a normal community, only those with carcinoma had a significant decrease in positive probability (P-value `r ifelse(all_probs_paired_wilcox["carcinoma", "BH_correction"] < 0.05, invisible("<"), invisible(">"))` 0.05); providing further evidence that treatment has the greatest effect in those with carcinoma.

5. CRC-associated bacterial populations are unique and they decline in relative abundance with treatment

6. Among the patients with carcinoma, the type of treatment did not significantly alter the amount of change to their bacterial communities.

Both chemotherapy and radiation did not provide any additional changes to the bacterial community in those treated for carcinoma (P-value `r ifelse(min(chemo_rad_summary$bh, na.rm = TRUE) < 0.05, invisible("<"), invisible(">"))` 0.05).

**Conclusions.** Our data partially supports the hypothesis that the microbiome changes after treatment towards a normal community. Individuals with carcinoma had more drastic differences to the overall community than those with adenoma. Common OTUs to all models were overwhelmingly from commensal bacteria, suggesting that these bacteria may be important in initial polyp formation, development of advanced adenoma, and transition to carcinoma.

@@

bacterial microbiome; colorectal cancer; polyps; FIT; post-surgery; risk factor. .

@@

Colorectal cancer (CRC) is the third most common cause of cancer deaths in the United States [@jemal_cancer_2010; @haggar_colorectal_2009]. The rate of disease mortality has seen a significant decrease, thanks mainly to improvements in screening [@jemal_cancer_2010]. Despite this improvement, there are still approximately 50,000 deaths from the disease per year [@haggar_colorectal_2009]. Current estimates indicate that 20-30% of those who undergo treatment will experience recurrence and 30-50% of those patients will die [@hellinger_reoperation_2006; @ryuk_predictive_2014]. Identification of methods to assess patients' risk of recurrence is of great importance to reduce mortality and healthcare costs.

[ 	fix: the following paragraph is mixing two things - mechanistic studies and biomarker studies. they need to be split up so that you establish that there is a mechanistic link and that we can use biomarkers to diagnose people with adenomas and carcinomas from those with normal colons. ]

Recent studies in humans and mouse models have demonstrated a connection between the  that altered membership and structure of the gut microbiome correlate with CRC pathogenesis [@zackular_manipulation_2016; @arthur_microbial_2014; @zackular_gut_2013]. Further, bacterial communities on the mucosa are altered between normal and tumor tissue [@dejea_microbiota_2014].
Collectively, these studies provide a tantalizing link between our gut bacteria and CRC and suggest the possibility of using biomarkers to diagnose disease.

Indeed, models created with microbiome data and clinical tests, such as Fecal Immunoglobulin Test (FIT), result in good predictions of CRC [@baxter_microbiota-based_2016; @zeller_potential_2014]. Although these studies show how changed microbial communities or invasion by inflammatory species can impact CRC progression [@flynn_metabolic_2016], they provide very little information as to whether treatment to remove lesions affects the presence of bacterial populations associated with those lesions.

@@

Using pre- (initial) and post- (follow up) treatment samples we tested the hypothesis that treatment causes detectable changes towards a normal microbiome in those with adenoma, advanced adenoma, and carcinoma. First, we assessed differences between samples pre- and post-treatment in adenoma, advanced adenoma, or carcinoma using alpha or beta diversity metrics. Second, we explored whether models built to classify adenoma, advanced adenoma, or carcinoma versus normal were able to identify specific community members that differed between pre- and post-treatment samples. We also used these models to assess whether changes in the community were toward a more normal microbiome. Finally, we assessed both whether surgery for adenomas and advanced adenomas provided larger community changes and whether chemotherapy or radiation provided additional changes to the microbiome over surgical resection. This study aims to provide evidence as to whether the altered CRC microbiome persists or shifts back towards normal after such interventions.

@@

***The Bacterial Community:***
Within our `r length(rownames(good_metaf))`-person cohort we tested whether the microbiome in patients with adenoma (N=`r change_summary["thetayc", "adn_n"]`), advanced adenoma (N=`r change_summary["thetayc", "srn_n"]`), or carcinoma (N=`r change_summary["thetayc", "crc_n"]`) had any broad differences between pre- and post-treatment samples. We found that carcinoma patients had a more dissimilar bacterial community between these samples than those with adenoma (P-value `r ifelse(change_summary["thetayc", "Pvalue_adnVcrc"] < 0.001, invisible("<"), invisible(">"))` 0.001) [Figure 1A]. Although no significant differences were observed between advanced adenoma and carcinoma there was an increase in the dissimilarity of pre- and post-treatment samples from adenoma (`r paste(format(change_summary["thetayc", "adn_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "adn_SD"], digits = 2))` (mean `r paste(label = "\u00B1")` SD)) to advanced adenoma (`r paste(format(change_summary["thetayc", "srn_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "srn_SD"], digits = 2))`) to carcinoma (`r paste(format(change_summary["thetayc", "crc_mean"], digits = 2), label = "\u00B1", format(change_summary["thetayc", "crc_SD"], digits = 2))`) [Figure 1A]. The bacterial community structure pre- post-treatment was visualized using NMDS for adenoma [Figure 1B] (PERMANOVA `r ifelse(beta_summary["Adenoma", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05), advanced adenoma [Figure 1C] (PERMANOVA `r ifelse(beta_summary["SRN", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05), and carcinoma [Figure 1D] (PERMANOVA `r ifelse(beta_summary["Carcinoma", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05). Interestingly, when pre- and post-treatment samples were compared, regardless of whether they were adenoma or carcinoma, there was no significant overall difference in beta diversity (PERMANOVA `r ifelse(beta_summary["All", "PERMANOVA"] < 0.05, invisible("<"), invisible(">"))` 0.05). There was no difference between pre- and post-treatment samples when investigating alpha diversity metrics for adenoma, advanced adenoma, or carcinoma [Table S1]. Additionally, there was also no difference in the relative abundance of any specific OTU between these samples for adenoma, advanced adenoma, or carcinoma only [Figure S1].

@@

***Common OTUs:***
Next, to identify which OTUs were important at all stages of disease, we identified common predictive OTUs within the adenoma, advanced adenoma, and carcinoma models. When we compared the three different models with each other there were a total of `r length(rownames(common_vars_summary))` common OTUs. Some of the most common taxonomic identifications belonged to *Bacteroides*, *Blautia*, *Anaerostipes*, *Lachnospiraceae*, and *Dorea*. The vast majority of the OTUs that were common between these models had classifications to bacteria typically thought of as commensal [Table S2]. The rank of these OTUs, based on median mean decrease in accuracy (MDA), varied higly depending on whether they were in the adenoma, advanced adenoma, or carcinoma model [Figure 3]. As an example, *Bacteroides* (OTU85) was ranked `r filter(crc_varsMDA, otu == "Otu000085") %>% select(rank)`^th^ in the carcinoma model but in both the adenoma and advanced adenoma model ranked `r filter(adn_varsMDA, otu == "Otu000085") %>% select(rank)`^th^ and `r filter(srn_varsMDA, otu == "Otu000085") %>% select(rank)`^st^ respectively [Figure 3]. .

@@

Many of the common OTUs that we identified taxonomically classified to potential butyrate producers (e.g. *Clostridiales*, *Roseburia*. and *Anaerostipes*) [Table S2]. Other OTUs classified to bacteria that are inhibited by polyphenols (e.g. *Bacteroides*). Both butyrate and polyphenols are thought to be protective against cancer, in part by reducing inflammation [@louis_gut_2014]. These protective compounds are derived from the breakdown of fiber, fruits, and vegetables by resident gut microbes. One example of this potential diet-microbiome-inflammation-polyp axis is that *Bacteroides*, which was highly prevalent in our models, are known to be increased in those with high non-meat based protein consumption [@zhu_intake_2016]. High protein consumption in general has been linked with an increased CRC risk [@mu_colonic_2016]. Conversely, *Bacteroides* are inhibited by polyphenols which are derived from fruits and vegetables [@ozdal_reciprocal_2016]. Our data fits with the hypothesis that the microbial metabolites from breakdown products within our own diet could not only help to shape the existing community but also have an effect on CRC risk and disease progression. Within this context the commensal community may be an important modifiable risk factor for monitoring and preventing CRC recurrence after treatment.

@@

Building off of our results, an area for future research, is that within our study we do not know whether individuals who were still classified as positive by the carcinoma model eventually had a subsequent CRC recurrence. This information would help to strengthen the case for this model keeping numerous individuals above the cutoff threshold even though at follow up they were diagnosed as no longer having carcinoma. It would also provide additional evidence that the microbiome could be used as a risk stratification tool in monitoring recurrence risk and whether different interventions could potentially change this community and lower the probability of future recurrence.

@@

**Figure 1: General Differences between Adenoma, Advanced Adenoma, and Carcinoma Groups After Treatment.*. A) A significant difference was found between the adenoma and carcinoma group for thetayc (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`). B) NMDS of the pre- and post-treatment samples for the adenoma group. C) NMDS of the pre- and post-treatment samples for the advanced adenoma group. D) NMDS of the pre- and post-treatment samples for the carcinoma group.

A: Convert mean to median w/ IQR. Add 0.00 to y-labels

@@

**Figure S2: ROC Curves of the Adenoma, Advanced Adenoma, and Carcinoma Models.*. A) Adenoma ROC curve: The light green shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark green line represents the model using 100% of the data set and what was used for subsequent classification. B) Advanced Adenoma ROC curve: The light yellow shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark yellow line represents the model using 100% of the data set and what was used for subsequent classification. C) Carcinoma ROC curve: The light red shaded areas represent the range of values of a 100 different 80/20 splits of the test set data and the dark red line represents the model using 100% of the data set and what was used for subsequent classification.

**Figure S3: Summary of Important Variables for the Adenoma, Advanced Adenoma, and Carcinoma Models.*. A) MDA of the most important variables in the adenoma model. The dark green point represents the mean and the lighter green points are the value of each of the 100 different runs. B) Summary of Important Variables in the advanced adenoma model. MDA of the most important variables in the SRN model. The dark yellow point represents the mean and the lighter yellow points are the value of each of the 100 different runs. C) MDA of the most important variables in the carcinoma model. The dark red point represents the mean and the lighter red points are the value of each of the 100 different runs.
